Evaluating machine translation for assimilation via a gap-filling task

### 2015-04-15 2015-eamt-assim/2015-eamt-assim.tex
back to 8 pages
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59824 72bbbca6-d526-0410-a7d9-f06f51895060

\section{Results and discussion}
\label{sec:results}

The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} 
show
the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
The evaluators' correct answer percentage is averaged over the number of evaluators. In addition to the percentage of correct answers we kept a record of the time taken to fill the gaps in one sentence. To reduce the noise from participants who were distracted during evaluation, when
calculating times we remove all the results over 6 minutes (the statistical mode is approximately
two minutes). The typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.

@@

\subsection{Annotator agreement}
After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.

@@

% a little on categories
\subsection{Results for different domains}
For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news. The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}. The MT system used in the evaluation has been targeted to translate texts from all three of the domains. Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories. Note that the source sentences paired with MT improve participants' performance in casual texts, compared to MT-only task mode. This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.

@@

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40\\
20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30\\
30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18\\
\hline
\end{tabular}
\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
\label{table:res-eus-spa}
\end{table*}

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
20\%&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
30\%&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
    \hline
  \end{tabular}
  \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
  \label{table:res-tat-rus}
\end{table*}

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.496&0.517&0.400&0.124\\
20\%&0.714&0.700&0.358&0.275\\
30\%&0.559&0.430&0.406&0.300\\

\hline
\end{tabular}
\caption {Basque--Spanish: Krippendorff Alpha measure} 
\label{table:alpha-eus-spa}
\end{table*}

\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.598&0.459 &0.711 &0.517 \\
20\%&0.740&0.667 &0.473 &0.261 \\
30\%&0.534&0.581 &0.411 &0.412 \\
\hline
\end{tabular}
\caption {Tatar--Russian: Krippendorff Alpha measure} 
\label{table:alpha-tat-rus}
\end{table*}

% results by text mode in tat-rus, with synonyms

\begin{table*}

\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
\hline
\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
\hline
\end{tabular}
\caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} 
\label{table:rus-casual} 
\end{table*}

\begin{table*}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
\hline
\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
\hline
\end{tabular}
\caption {Tatar--Russian, legal texts: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-legal} 
\end{table*}

\begin{table*}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
\hline
\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
\hline
\end{tabular}
\caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} 
\label{table:rus-news} 
\end{table*}

>>>>

\section{Results and discussion}
\label{sec:results}
\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Basque--Spanish} & \multicolumn{4}{|c|}{Tatar--Russian} 
\\\cline{2-9}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}& \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} 
\\
\hline
10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40
&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30
&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18
&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
\hline
\end{tabular}
\caption{Average number of gaps successfully filled (\%), using a synonym list} 
\label{table:res}
\end{table*}

The results are presented in this section. Table \ref{table:res} 
shows
the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
The evaluators' correct answer percentage is averaged over the number of evaluators. In addition to the percentage of correct answers we kept a record of the time taken to fill the gaps in one sentence. To reduce the noise from participants who were distracted during evaluation, when
calculating times we remove all the results over 6 minutes (the statistical mode is approximately
two minutes). The typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.

@@

\subsection{Annotator agreement}
After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Table~\ref{table:alpha}. 
\begin{table*}
\centering
  \begin{tabular}{|l|r|r|r|r|r|r|r|r|}
  \hline
 \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Basque--Spanish}&\multicolumn{4}{|c|}{Tatar--Russian} \\\cline{2-9}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}& \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
10\%&0.496&0.517&0.400&0.124&0.598&0.459 &0.711 &0.517\\
20\%&0.714&0.700&0.358&0.275&0.740&0.667 &0.473 &0.261\\
30\%&0.559&0.430&0.406&0.300&0.534&0.581 &0.411 &0.412\\

\hline
\end{tabular}
\caption{Krippendorff Alpha measure} 
\label{table:alpha}
\end{table*}

@@

% a little on categories
\subsection{Results for different domains}
\begin{table*}
\centering
\begin{tabular}{|r|r |*{4}{c}|}
\hline
\multicolumn{2}{|c|}{} &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
\hline
\textbf{Domain} & \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
\hline
 \multirow{3}{*}{\textbf{Casual}} &
10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
& 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
& 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
\hline
 \multirow{3}{*}{\textbf{Legal}} &
10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
&20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
&30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
\hline
 \multirow{3}{*}{\textbf{News}} &
10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
&20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
&30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
\hline

\end{tabular}
\caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains.} 
\label{table:rus} 
\end{table*}

For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news. The results by domains are displayed in table~\ref{table:rus}. The MT system used in the evaluation has been targeted to translate texts from all three of the domains. Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories. Note that the source sentences paired with MT improve participants' performance in casual texts, compared to MT-only task mode. This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.

@@

% \begin{table*}
% \centering
%   \begin{tabular}{|l|r|r|r|r|}
%   \hline
%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40\\
% 20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30\\
% 30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18\\
% \hline
% \end{tabular}
% \caption{Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
% \label{table:res-eus-spa}
% \end{table*}

% \begin{table*}
% \centering
%   \begin{tabular}{|l|r|r|r|r|}
%     \hline
%     \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
% 20\%&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
% 30\%&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
%     \hline
%   \end{tabular}
%   \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
%   \label{table:res-tat-rus}
% \end{table*}

% \begin{table*}
% \centering
%   \begin{tabular}{|l|r|r|r|r|}
%   \hline
%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&0.496&0.517&0.400&0.124\\
% 20\%&0.714&0.700&0.358&0.275\\
% 30\%&0.559&0.430&0.406&0.300\\

% \hline
% \end{tabular}
% \caption {Basque--Spanish: Krippendorff Alpha measure} 
% \label{table:alpha-eus-spa}
% \end{table*}

% \begin{table*}
% \centering
%   \begin{tabular}{|l|r|r|r|r|}
%   \hline
%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&0.598&0.459 &0.711 &0.517 \\
% 20\%&0.740&0.667 &0.473 &0.261 \\
% 30\%&0.534&0.581 &0.411 &0.412 \\
% \hline
% \end{tabular}
% \caption {Tatar--Russian: Krippendorff Alpha measure} 
% \label{table:alpha-tat-rus}
% \end{table*}

% results by text mode in tat-rus, with synonyms
% \begin{table*}
% \centering
% \begin{tabular}{|r|r |*{4}{c}|}
% \hline
%   &&\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
% \hline
% & \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
%  \multirow{3}{*}{\textbf{Casual}} &
% 10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
% & 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
% & 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
% \hline
%  \multirow{3}{*}{\textbf{Legal}} &
% 10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
% &20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
% &30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
% \hline
%  \multirow{3}{*}{\textbf{News}} &
% 10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
% &20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
% &30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
% \hline

% \end{tabular}
% \caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains.} 
% \label{table:rus} 
% \end{table*}

% \begin{table*}

% \centering
% \begin{tabular}{|r |*{4}{c}|}
% \hline
%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
% \hline
% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
% 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
% 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
% \hline
% \end{tabular}
% \caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} 
% \label{table:rus-casual} 
% \end{table*}

% \begin{table*}
% \centering
% \begin{tabular}{|r |*{4}{c}|}
% \hline
%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
% \hline
% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
% 20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
% 30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
% \hline
% \end{tabular}
% \caption {Tatar--Russian, legal texts: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-legal} 
% \end{table*}

% \begin{table*}
% \centering
% \begin{tabular}{|r |*{4}{c}|}
% \hline
%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
% \hline
% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
% \hline
% 10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
% 20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
% 30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
% \hline
% \end{tabular}
% \caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} 
% \label{table:rus-news} 
% \end{table*}
